### 2021-04-27 eco_footprint.Rmd
download ecological footprint data

---
title: "Ecological footprint"
author: "Juan Rocha"
date: "4/26/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(patchwork)

theme_set(theme_light())
```

A short document to remind myself how to get ecological footpint data. The global footprint network has an API documented [here](https://data.footprintnetwork.org/#/api). Through their service one can download all the time series of the ecological foot print per country. The API uses `curl`, so one can do it from `bash` directly or using `Rcurl`.

Here is an example of the command I used on bash (with the key from email):

```{bash}
# curl -u juan:<api-key-here> http://api.footprintnetwork.org/v1/years -o years.json
```

That way I got the `types`, `countries` and `years` json files. Now I parse them into data frames to extract the relevant variables:

```{r}
countries <- jsonlite::read_json("data/countries.json", simplifyVector = TRUE) %>% 
    as_tibble()
years <- jsonlite::read_json("data/years.json", simplifyVector = TRUE) %>% 
    as_tibble()
types <- jsonlite::read_json("data/types.json", simplifyVector = TRUE) %>% 
    as_tibble()

```

```{bash}
#curl -u juan:<api-key-here>  "https://api.footprintnetwork.org/v1/data/all/[1961-2017]" -o "all_#1.json"

```

```{r}
test <- jsonlite::read_json("data/ecofootprint/all_2000.json", simplifyVector = TRUE) %>% 
    as_tibble()
```

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()
```

```{r}
ecofp
```

```{r}
ecofp %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = shortName), show.legend = FALSE) +
    facet_wrap(~record, scales = "free_y" ) +
    theme_light(base_size = 9)
```