### 2021-05-25 missing_values.Rmd

---
title: "Missing values"
author: "Juan Rocha"
date: "May 2021"
output:
  html_document:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
    self_contained: true

  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

theme_set(theme_light(10))
```

## Ecological footprint

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()

ecofp
```

```{r}
ecofp %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = shortName), show.legend = FALSE) +
    facet_wrap(~record, scales = "free_y" ) 
```

```{r}
ecofp %>%
    filter(record == "EFConsPerCap") %>%
    mutate(missing = is.na(value)) %>% 
    ggplot(aes(year, countryName)) +
    geom_tile(aes(fill = value)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```

```{r fig.height=3, fig.width=8}
g1 <- ecofp %>%
    filter(record == "EFConsPerCap") %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(year) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    ggplot(aes(year, n_nas)) + 
    geom_col(aes(fill = (year>1995)), alpha = 0.5) + 
    coord_flip() +
    theme_light(base_size = 7) +
    theme(legend.position = c(0.8,0.8)) + labs(title = "Missing values per year")

g2 <- ecofp %>%
    filter(record == "EFConsPerCap", year > 1995) %>%
    select(countryName, year, value) %>% 
    pivot_wider(id_col = countryName, names_from = year, values_from = value) %>% 
    pivot_longer(cols = 2:last_col(), names_to = "year", values_to = "value") %>% 
    mutate(missing = is.na(value), year = as.numeric(year)) %>% 
    group_by(countryName) %>% 
    summarise(n_nas  = sum(missing)) %>% 
    filter(n_nas > 0) %>% 
    mutate(countryName = as_factor(countryName) %>% fct_reorder(., n_nas)) %>% 
    ggplot(aes(countryName, n_nas)) + 
    geom_col(aes(fill = (n_nas < 11)), alpha = 0.5) + coord_flip() +
    theme_light(base_size = 7) + 
    labs(title = "Missing values per country after 1995") +
    theme(legend.position = c(0.8, 0.2))

g1 + g2
```

An alternative is use the data after 1995 where `NA`s start decreasing but still give us a couple of decades worth of data to work with. Another is limiting to countries where the count of `NA`s is less than 10.

## Gini: Inequality

```{r}
inq <- readxl::read_xlsx(
    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
inq
```
Need to read the docs carefully, it seems many measurements per year and it's hard to tell what is the unit per time that we should use for the analysis.

```{r}
inq %>% 
    ggplot(aes(year, gini_reported)) +
    geom_line(aes(color = country), show.legend = FALSE) +
    facet_grid(region_un ~ resource)
```

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    ggplot(aes(gini_reported, palma)) +
    geom_miss_point(size = 1, alpha = 0.5) +
    theme(legend.position = c(0.1,0.8))
```

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    filter(country == "Australia")
```
Note there are multiple estimates per year-country because the database report different sources.

```{r}
inq %>% 
    filter(year > 1950, resource_detailed == "Income, gross") %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini") %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
```{r}
inq %>% 
    filter(!is.na(palma), year > 1990) %>% select(country, year, palma) %>% 
    group_by(country, year) %>% 
    summarize(palma_mean = mean(palma, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = palma_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "palma_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = palma_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
To be fair with Gini, let's to the same over all Gini's reported. Even if Gini comes computed by different studies and in different currencies, the coefficient is always between 0-100, meaning it is comparable.

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year)) %>%
    ggplot(aes(year, country)) +
    geom_tile(aes(fill = gini_mean)) +
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.y = element_text(size = 4), legend.position = "top")
```
So how many country / years can we use? Remember the data has `r inq$country %>% unique() %>% length()` countries

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(year) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / 200)
```
We are dealing with ~50% of missing values on the best years, at best we can work with 100 countries. Which ones?

```{r}
inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarise(n_nas = sum(missing), missingness = n_nas / (2018-1990)) %>% 
    arrange(missingness)
```

If we allow for 60% missing values, we get 80 countries, 30% NAs is ~60 countries. If needed one can impute `NA`s with a mean trend. But how much is that messing with the real trends? Double click on the legend to select particular countries:

```{r fig.width=8}
g <- inq %>% 
    filter(!is.na(gini_reported), year > 1990) %>% select(country, year, gini_reported) %>% 
    group_by(country, year) %>% 
    summarize(gini_mean = mean(gini_reported, na.rm = TRUE)) %>% 
    pivot_wider(id_cols = country, names_from = year, values_from = gini_mean) %>% 
    pivot_longer(2:last_col(), names_to = "year", values_to = "gini_mean") %>%
    mutate(year = as.numeric(year), missing = is.na(gini_mean)) %>% 
    ggplot(aes(year, gini_mean)) +
    geom_line(aes(color = country), show.legend = FALSE, size = 0.5) 
ggplotly(g, tooltip = "country") 
```

This justifies why the step-wise approach that Tong first proposed makes sense. If we restrict the analysis to good-data countries, we miss key actors like South Africa, the country with highest inequality in the world.

## GNI: Prosperity

```{r}
dat1 <- read_csv(
    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
    janitor::clean_names()
```

```{r}
dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(color = country_name), show.legend = FALSE) +
    #geom_miss_point(show.legend = FALSE) +
    facet_wrap(~indicator_name, scales = "free_y")
```

```{r}
## How many missing values? 10506, but who has more?
missingness <- dat1 %>% 
    select(-x33) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(ismissing = is.na(value)) %>% 
    group_by(country_code, indicator_name, country_name) %>% 
    summarize(missingness = sum(ismissing)/n())

missingness %>% 
    ggplot(aes(country_code, indicator_name)) +
    geom_tile(aes(fill = missingness)) + 
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.x = element_text(angle =90), legend.position = "top")
```

Decide who to exclude and which variable to use

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(indicator_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(x)
```

```{r}
missingness %>% 
    ungroup() %>% 
    group_by(country_name) %>% 
    summarize(x = mean(missingness)) %>% 
    arrange(desc(x)) 
```
It seems that `GNI per capita (current US$)` is the one with less missing values. Tong recomments using the one with Atlas method wich is close on missingness, only 18%. A threshold of 50% missigness per country will left out places like Somalia, Iraq, Costa Rica, Zambia, China or Chile. Perhaps 60% is tolerable?

## Analysis

The bottle neck is centaintly our proxies of inequality, no matter what we use. Below a mock up example with some arbitrary choices.

```{r}

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1990)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1990) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

  )

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-x33) %>% 
      filter(indicator_name == "GNI per capita (current US$)") %>% 
      pivot_longer(cols = x1990:x2017, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1990) %>% 
      rename(country = country_name, gni = value)
  )

dat

```
Each path is a country trajectory over time for the data we have available since 1990. Here are the three phase space graphs, one for each combination of variables in the trilemma. Ideally, we want to understand if being in a particular corner of the space force you to certain trajectories in other dimensions (e.g. being poor country forces you to treat the environment badly while you reach a minimum of equality or economic activity). The red dots are missing values and remind us why it is difficult to perform a full dynamic analysis, and many dots that are not missing are left alone, mining it will be hard to reconstruct time series for some countries.

```{r}
p1 <- dat %>% 
  ggplot(aes(EFconsPerCap, gini_mean, group = country)) +
  geom_miss_point(size = 0.5) +
  geom_path(size = 0.2, color = "gray50") +
  theme(legend.position = c(0.8, 0.8))

p2<- dat %>% 
  ggplot(aes(gni, gini_mean, group = country)) +
  geom_miss_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, color = "gray50") +
  scale_x_log10()

p3<- dat %>% 
  ggplot(aes(gni, EFconsPerCap, group = country)) +
  geom_miss_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, color = "gray50") +
  scale_x_log10()

p1 + p2 + p3
```

```{r}
# g <- dat %>% 
#   ggplot(aes(EFconsPerCap, gini_mean, group = country)) +
#   geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
#   geom_path(aes(color = country), size = 0.3) +
#   theme_light(base_size = 7)
# 
# ggplotly(g, tooltip = c("country", "gini_mean", "EFconsPerCap"),
#          width = 640, height = 480)
```

```{r}
g <- dat %>% 
  ggplot(aes(gni, gini_mean, group = country)) +
  geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
  geom_path(aes(color = country), size = 0.3) +
  scale_x_log10() +
  theme_light(base_size = 7)

ggplotly(g)
```

```{r}
g <- dat %>% 
  ggplot(aes(gni, EFconsPerCap, group = country)) +
  geom_point(aes(color = country), size = 0.5, alpha = 0.5) +
  geom_path(aes(color = country), size = 0.3) +
  scale_x_log10()+
  theme_light(base_size = 7)

ggplotly(g)
```