### 2022-02-11 Trilemma_clustering.Rmd
improved figures for paper

---
title: "Trilemma"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook: 
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)
# library(ggbiplot)

#theme_set(theme_light(base_size = 12))
```

## Data

```{r}
ecofp <- list()
fls <- list.files("data/ecofootprint/")
ecofp <- map(fls, function(x) {
    jsonlite::read_json(
        path = paste0("data/ecofootprint/", x),
        simplifyVector = TRUE)
    }) %>% 
    bind_rows()

ecofp
```

```{r}
inq <- readxl::read_xlsx(
    path = "~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/WIID_06MAY2020.xlsx")
inq
```

```{r}
dat1 <- read_csv(
    "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") %>% 
    janitor::clean_names()

missingness <- dat1 %>% 
    select(-last_col()) %>% 
    filter(str_detect(indicator_name, "GNI")) %>% 
    pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>% 
    mutate(year = str_remove(year, "x")) %>% 
    mutate(year = as.numeric(year)) %>% 
    mutate(ismissing = is.na(value)) %>% 
    group_by(country_code, indicator_name, country_name) %>% 
    summarize(missingness = sum(ismissing)/n())

missingness%>% 
    ggplot(aes(country_code, indicator_name)) +
    geom_tile(aes(fill = missingness)) + 
    scale_fill_viridis_c(option = "C") +
    theme(axis.text.x = element_text(angle =90), legend.position = "top")
```

## Analysis

The bottle neck is certaintly our proxies of inequality, no matter what we use. Based on the missing data explorations, inequality data improves from 1995 onward.

```{r}

dat <- ecofp %>%
  filter(record == "EFConsPerCap") %>%
  select(countryName, isoa2, year, value) %>% 
  rename(EFconsPerCap = value, country = countryName) %>% 
  filter(year >= 1995)

dat <- dat %>% 
  left_join(
    .,
    inq %>% 
      filter(!is.na(gini_reported), year >= 1995) %>% 
      select(country, isoa2 = c2, country_code = c3, year, gini_reported) %>% 
      group_by(country, year, isoa2, country_code) %>% 
      summarize(gini_mean = mean(gini_reported, na.rm = TRUE))

  )

dat <- dat %>% 
  left_join(
    ., 
    dat1 %>% 
      select(-last_col()) %>% 
      filter(indicator_name == "GNI per capita (US$)") %>% 
      pivot_longer(cols = x1990:x2019, names_to = "year", values_to = "value") %>%
      mutate(year = str_remove(year, "x")) %>% 
      mutate(year = as.numeric(year)) %>% 
      mutate(ismissing = is.na(value)) %>% 
      filter(year >= 1995) %>% 
      rename(country = country_name, gni = value)
  )

dat

```

Compute missing values for inequality:
```{r}
missingness <- dat %>% 
    mutate(ismissing = is.na(gini_mean)) %>% 
    group_by(country) %>% 
    summarize(missingness = sum(ismissing)/n()) %>% 
    arrange(missingness)

key_countries <- missingness %>% 
    filter(missingness <= 0.3) %>% 
    pull(country)

droped_countries <- missingness %>% 
    filter(missingness > 0.3) %>% 
    pull(country)
```

```{r}
#library(tsibbledata)
df_dat <- dat %>% 
    select(-starts_with("indicator")) %>% 
    filter(country %in% key_countries) %>% 
    as_tsibble(key = country, index = year)

df_dat <- df_dat %>% 
    fill_gaps(.full = TRUE)

```

```{r}

df_dat <- df_dat %>% 
    group_by(country) %>% 
    mutate(
        gini = na_interpolation(gini_mean),
        gni2 = na_interpolation(gni),
        EFconsPerCap = na_interpolation(EFconsPerCap)
    ) 
```

```{r message=FALSE, warning=FALSE}
p <- df_dat %>% 
    select(year, `Ecological footprint` = EFconsPerCap, GNI = gni2, Gini = gini) %>% 
    GGally::ggpairs(
        columns = 3:5 ,
    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
    # upper = list(continuous = GGally::wrap("cor", size = 2))
    ) + theme_light(base_size = 6)

# ggsave(
#     plot = p,
#     file = "correlogram.png", path = "figures/", device = "png",
#     width = 4, height = 4, dpi = 300, bg = "white"
# )
p + theme_light(base_size = 20)
```

```{r}
df_dat <- df_dat %>% 
    select(EFconsPerCap, gini, gni2) %>% 
    filter(!is.na(gni2))   # lost 2 more countries due to NAs

```

```{r}
df_dat %>% 
    pivot_longer(cols = c("EFconsPerCap", "gini", "gni2"),
                 names_to = "variable", values_to = "value") %>% 
    ggplot(aes(year, value)) +
    geom_line(aes(group = country)) +
    facet_wrap(~variable, ncol = 1, scales = "free") +
    theme_light(base_size = 15)
```

```{r}
p1 <- df_dat %>% 
    ggplot(aes(EFconsPerCap, gini, group = country)) +
    geom_point(size = 0.5) +
    geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 0, xmax = 1.68, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom ="rect",xmin = 0, xmax = 1.68, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect", xmin = 1.68, xmax = 18, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom="rect", xmin = 1.68, xmax = 18, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
    labs(y = "Gini", x = "Ecological footprint", tag = "A") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p2<- df_dat %>% 
  ggplot(aes(gni2, gini, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year)) +
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 20, ymax = 30,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 30, ymax = 62,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 20, ymax = 30,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 30, ymax = 62,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Gini", x = "GNI [log]", tag = "B") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

p3<- df_dat %>% 
  ggplot(aes(gni2, EFconsPerCap, group = country)) +
  geom_point(size = 0.5, show.legend = FALSE) +
  geom_path(size = 0.2, aes(color = year))+
    annotate(geom = "rect", xmin = 10, xmax = 12746, ymin = 0, ymax = 1.68,
             fill = "orange", alpha = 0.2) +
    annotate(geom ="rect",xmin = 10, xmax = 12746, ymin = 1.68, ymax = 18,
             fill = "red", alpha = 0.2) +
    annotate(geom ="rect", xmin = 12746, xmax = 105000, ymin = 0, ymax = 1.68,
             fill = "skyblue", alpha = 0.2) +
    annotate(geom="rect", xmin = 12746, xmax = 105000, ymin = 1.68, ymax = 18,
             fill = "orange", alpha = 0.2) +
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 12746, linetype = 2, color = "gray50") +
    scico::scale_color_scico("Year", palette = "bilbao") +
      scale_x_log10() +
    labs(y = "Ecological footprint", x = "GNI [log]", tag = "C") + theme_light() +
    theme(legend.position = "top", panel.grid = element_blank(),
          legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm"))

# ggsave(
#     plot = p1 + p2 + p3 + plot_layout(guides = "collect") & 
#         theme_light(base_size = 7) & 
#         theme(legend.position = "bottom", panel.grid = element_blank(),
#           legend.key.width = unit(10, "mm"), legend.key.height = unit(3,"mm")),
#     path = "figures/", file = "cube_time.png", device = "png", dpi = 300,
#     width = 7, height = 3, bg = "white"
# )

p1 + p2 + p3 + 
    plot_layout(guides = "collect") & theme_light(base_size = 18) & 
    theme(legend.position = "bottom", panel.grid = element_blank(),
          legend.key.width = unit(20, "mm"), legend.key.height = unit(3,"mm"))
```

```{r}
df_dat %>% 
    as_tibble() %>% 
    group_by(country) %>% 
    summarize(
        mean_ef = mean(EFconsPerCap),
        mean_gni = mean(gni2),
        mean_gini = mean(gini),
        var_ef = var(EFconsPerCap),
        var_gni = var(gni2),
        var_gini = var(gini)
    ) %>% 
    arrange(desc(mean_ef))
```
### Experimenting with phase plane

```{r}
df_dat <- df_dat |> 
    # change scale of GNI to log so the magnitud is not so dominated by $
    mutate(gni_log = log(gni2)) |> 
    mutate(
        d_EF = slide_dbl(EFconsPerCap, diff, .before = 1, .after = 0, .complete = TRUE),
        d_gini = slide_dbl(gini, diff, .before = 1, .after = 0, .complete = TRUE),
        d_gni = slide_dbl(gni_log, diff, .before = 1, .after = 0, .complete = TRUE)
    ) |> 
    mutate(
        magni = sqrt(d_EF^2 + d_gini^2 + d_gni^2), 
        theta = atan(d_gini/d_EF),
        gamma = atan(d_gini/d_gni)
    ) 
```

```{r}
df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    summarize(mgt = sum(magni, na.rm = TRUE), 
              theta = sum(theta, na.rm = TRUE),
              gamma = sum(gamma, na.rm = TRUE)) |> 
    arrange(desc(mgt))
```
```{r fig.width=6, fig.height=3}
p <- df_dat |> 
    as_tibble() |> 
    select(country, year, d_EF, d_gini, d_gni) |> 
    pivot_longer(cols = starts_with("d_"), names_to = "vars", values_to = "vals") |> 
    filter(!is.na(vals)) |> 
    mutate(vars = case_when(
        vars == "d_EF" ~ "Ecological Footprint",
        vars == "d_gni" ~ "GNI [log]",
        vars == "d_gini" ~ "Gini"
    )) |> 
    ggplot(aes(year, vals, group = country)) +
    geom_path(size = 0.2, alpha = 0.5) +
    facet_wrap(~vars, scales = "free") +
    labs(x = "Year", y = "First difference") +
    theme_light(base_size = 7)

# ggsave(
#     plot = p, path = "figures/", file = "first_diff.png", device = "png",
#     width = 6, height = 3, dpi = 300, bg = "white")
p + theme_light(base_size = 20)
```
```{r fig.width=3, fig.height=3}
c1 <- df_dat |> 
    ggplot(aes(EFconsPerCap, gini)) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = 1.68, linetype = 2, color = "gray50") +
    geom_density2d() +
    geom_point(aes(color = gni_log)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico("GNI [log]") +
    labs(y = "Gini", x = "Ecological footprint") +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

c1 + theme_light(base_size = 18)
```
```{r fig.width=3, fig.height=3}
c2 <- df_dat |> 
    ggplot(aes(gni2, gini)) +
    geom_density2d() +
    geom_point(aes(color = EFconsPerCap)) +
    geom_path(aes(group = country), size = 0.15) +
    geom_hline(yintercept = 30, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    scico::scale_color_scico("Ecological footprint",palette = "berlin") +
    labs(x= "GNI [log]", y = "Gini") +
    scale_x_log10() +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

c2 + theme_light(base_size = 7)
```
```{r fig.width=3, fig.height=3}
c3 <- df_dat |> 
    ggplot(aes(gni2, EFconsPerCap)) +
    geom_density2d() +
    geom_point(aes(color = gini)) +
    geom_path(aes(group = country), size = 0.15) +
    scico::scale_color_scico("Gini") +
    labs(x = "GNI", y = "Ecological footprint") +
    scale_x_log10()+
    geom_hline(yintercept = 1.68, linetype = 2, color = "gray50") +
    geom_vline(xintercept = (12546), linetype = 2, color = "gray50") +
    theme_light(base_size = 7) +
    theme(legend.position = "top", panel.grid = element_blank())

c3 + theme_light(base_size = 18)
```
Who is the weird trajectory out there, world? No it's Luxemburg

```{r}
df_dat |> 
    arrange(desc(EFconsPerCap))
```

## Clustering

I'm not sure if the clustering requires one column per variable * year. Since really the unit of observation is country. Need to read upon it.

```{r fig.width=6, fig.height=2}
df_summary <- df_dat |> 
    as_tibble() |> 
    ungroup() |> 
    group_by(country) |> 
    dplyr::summarize(sd_magnitude = sd(magni, na.rm = TRUE),
              sd_theta = sd(theta, na.rm = TRUE),
              sd_gamma = sd(gamma, na.rm = TRUE), 
              EF = mean(EFconsPerCap, na.rm = TRUE),
              Gini = mean(gini, na.rm = TRUE),
              GNI = mean(gni_log, na.rm = TRUE),
              sd_EF = sd(EFconsPerCap, na.rm = TRUE),
              sd_Gini = sd(gini, na.rm = TRUE),
              sd_GNI = sd(gni_log, na.rm = TRUE),
              magnitude_m = mean(magni, na.rm = TRUE), 
              theta_m = mean(theta, na.rm = TRUE),
              gamma_m = mean(gamma, na.rm = TRUE),
              magnitude_a = sum(magni, na.rm = TRUE), 
              theta_a = sum(theta, na.rm = TRUE),
              gamma_a = sum(gamma, na.rm = TRUE))

m <- "ward.D2" # minimize the total within-cluster variance

clust_num <- NbClust::NbClust( 
    data = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)),
    #distance = "maximum",
    min.nc = 2, max.nc = 7, 
    method = m, 
    index = 'all')
```

The Hubert index (which is graphical) favours 4 groups while D index favours 3 groups. Thus in total 9 metrics proposed 2 groups, 9 metrics proposed 3 groups, and all other options are minorities. Let's continue the analysis with 3 to favour diversity.

```{r}
## Stability & Internal validation
stab <- clValid::clValid(
    obj = df_summary %>% as.data.frame() %>% 
        select(where(is.numeric)) %>%
        as.matrix(),
    nClust = c(2:7),
    clMethods = c("hierarchical", "kmeans",  #"som", "model", "diana", "fanny",
                 "sota", "pam", "clara", "agnes"),
    validation = c('stability', "internal"),
    #metric = "manhattan",
    method = "ward.D2",
    verbose = FALSE
) ## Hierarchical is the optimal method

clValid::optimalScores(stab)
```
The stability analysis also supports the choice of either 2 or 7 clusters. Partition around medioids does perform well with 9 clusters, but others are supporting the 2 clusters.

```{r}
stab@clusterObjs$pam$`2`$clustering
```

```{r fig.width=7, fig.height=8}
df_summary |> 
    mutate(clus = stab@clusterObjs$pam$`2`$clustering) |> 
    mutate(country = as_factor(country)) |> 
    mutate(country = fct_reorder2(country, magnitude, Gini)) |> 
    pivot_longer(cols = sd_magnitude:gamma, names_to = "vars", values_to = "vals") |> 
    ggplot(aes(vals, country)) +
    geom_col(aes(fill = as.factor(clus))) +
    facet_grid(clus~vars, scales = "free") +
    theme_light(base_size = 10)
```

```{r}
df_summary |>
    select(-starts_with("country")) |> 
    GGally::ggpairs()
```

```{r fig.width=3, fig.height=3}
pc <- prcomp(df_summary |> 
                 select(-starts_with("country")),
             center = TRUE, scale. = TRUE)
## Fix short names for plotting
df_summary <- df_summary |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique())

pc1 <- ggbiplot::ggbiplot(pc, groups = as.character(stab@clusterObjs$pam$`2`$clustering),
         ellipse = TRUE, labels = df_summary$country_code, labels.size = 2.5,
         varname.size = 4) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    theme_light(base_size = 18) +
    theme(legend.position = c(0.2, 0.2), panel.grid = element_blank())

pc1
```

Note to self: It does not make sense to plot the average distance traversed in the space per country. Take Luxemburg as example, it started and ended verly close by. The absolute distance is what I'm calculating with the euclidian distance on that plane, then the first-diff values are just increases or decreases with respect to initial conditions. Averages are then misleading, because on average Luxemburg moved a lot (in circles).

Averages are worth calculating for the parameter space, the vector field of the cube. But not for individual countries, or particules moving within the cube. Be aware of that when presenting and interpreting the results. 
```{r fig.height=3,fig.width=3}
a1 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(EF, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = EF, yend = Gini, color = as.factor(Group)),
        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
    geom_text(
        aes(x = EF *1.1, y = Gini * 1.1, label = country_code, colour = Group), size = 2) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    labs(x = "Ecological Footprint", y = "Gini") +
    theme_light(base_size = 6) +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))

a1

c1 + annotation_custom(
        grob = ggplotGrob(a1),
        xmin = 10, xmax = 18, ymin = 40, ymax = 60
    ) 
```

```{r fig.width=2, fig.height=2}
a2 <- df_dat |> 
    as_tibble() |> 
    group_by(country) |> 
    dplyr::summarize(
        EF = sum(d_EF, na.rm = TRUE),
        Gini = sum(d_gini, na.rm = TRUE),
        GNI = sum(d_gni, na.rm = TRUE)
    ) |> 
    mutate(Group = as.character(stab@clusterObjs$pam$`2`$clustering)) |> 
    left_join(
        dat |> select(country, country_code) |> 
            filter(!is.na(country_code)) |> unique()) |> 
    ggplot(aes(GNI, Gini)) +
    geom_segment(
        aes(x = 0, y = 0, xend = GNI, yend = Gini, 
            color = as.factor(Group)), 
        arrow = arrow(length = unit(0.15, "cm")), size = 0.15) +
    geom_text(aes(x = GNI * 1.1, y = Gini *1.1, 
                  label = country_code, colour = Group), size = 2) +
    scale_color_manual("Groups", values = c("orange", "purple")) +
    theme_light(base_size = 6) +
    theme(panel.grid = element_blank(), legend.position = c(0.8,0.2))

c2 + annotation_custom(
        grob = ggplotGrob(a2),
        xmin = 1e+01, xmax = 1e+03, ymin = 40, ymax = 60
    )

a2
```

### Old stuff

Here is an attempt with another package `kml` that take into consideration the fact that our data is longitudinal. However, looking into the documentation, it seems that it only handles a variable at the time. For the sake of testing:

```{r}
# library(kml)

test_dat <- df_dat %>% 
  select(country, gini, year) %>% 
  as_tibble() %>% 
  pivot_wider(names_from = year, values_from = gini, names_prefix = "yr_") %>% 
  as_data_frame() %>% 
  kml::cld(idAll = "country", timeInData = 2:29)

```
An error I don't understand.