### 2022-03-02 SDGs_UN.Rmd
data explorations for SDGs datasets

---
title: "SDGs UN data"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

This notebooks explores synergies and trade-offs between inequality and other SDGs with the UN dataset on sustainable development goals at the national scale.

```{r}
dat <- read_csv(
    file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> 
    janitor::clean_names()

```

Filter out non-countries:

```{r}
# dat |> pull(geo_area_name) |> unique()

no_country <- c(
    "World", "South America" ,"Oceania","Western Africa","Central America", "Eastern Africa", 
    "Northern Africa","Middle Africa", "Southern Africa","Northern America", "Caribbean" ,"Eastern Asia", 
    "Southern Asia" ,"South-Eastern Asia", "Southern Europe", "Australia and New Zealand",  "Melanesia", 
    "Polynesia" , "Central and Southern Asia" ,"Caucasus and Central Asia" , "Asia" ,"Central Asia" , 
    "Western Asia" , "Europe", "Eastern Europe", "Northern Europe" , "Western Europe",
    "Least Developed Countries (LDCs)","Sub-Saharan Africa" , "Latin America and the Caribbean" ,
    "Landlocked developing countries (LLDCs)", 
    "Western Asia (exc. Armenia, Azerbaijan, Cyprus, Israel and Georgia)", "Europe and Northern America",
    "Developed regions (Europe, Cyprus, Israel, Northern America, Japan, Australia & New Zealand)", 
    "Eastern Asia (excluding Japan)","Oceania (exc. Australia and New Zealand)", 
    "Micronesia (Federated States of)" , "Small island developing States (SIDS)" ,
    "Sub-Saharan Africa (inc. Sudan)", "Northern Africa (exc. Sudan)","Northern Africa and Western Asia", 
    "Eastern and South-Eastern Asia" , "French Polynesia" , "China, Macao Special Administrative Region"  , 
    "New Caledonia" ,"British Virgin Islands" ,"American Samoa", "Guam" ,"Northern Mariana Islands",
    "Tokelau", "Wallis and Futuna Islands", "Holy See",  "China, Hong Kong Special Administrative Region",
    "Africa", "Americas", "Micronesia", "Southern Asia (excluding India)" ,
    "Eastern Asia (excluding Japan and China)" , "Developing regions" , "Other non-specified areas",
    "Faroe Islands",  "Falkland Islands (Malvinas)","Greenland" , "Guadeloupe", "Martinique", "Montserrat",
    "Réunion" ,"United States Virgin Islands" , "International Centers (FAO)" , "Regional Centres (FAO)",
    "Mayotte" , "Cayman Islands" , "Netherlands Antilles","Curaçao" , "Sint Maarten (Dutch part)",
    "Bonaire, Sint Eustatius and Saba" , "Gibraltar" , "Saint Pierre and Miquelon" , "Channel Islands",
    "British Indian Ocean Territory" , "Christmas Island" , "Cocos (Keeling) Islands",
    "South Georgia and the South Sandwich Islands" , "Åland Islands" , "French Southern Territories",
    "Heard Island and McDonald Islands" , "Norfolk Island", "United States Minor Outlying Islands" ,
    "Pitcairn",  "Saint Barthélemy" , "Saint Helena", "Saint Martin (French Part)" , "Western Sahara",
    "Svalbard and Jan Mayen Islands","Guernsey", "Jersey" ,"European Union", "Bouvet Island", "Antarctica" ,
    "Iraq (Central Iraq)" , "Iraq (Kurdistan Region)" , "United Kingdom (England and Wales)" ,
    "European Union (EU) Institutions" 
)

# Included Northern Ireland and Scotland as countries although they are listed as part of the UK
## Dataset reduced to countries and SDGs 10, 14, 15
dat <- dat |> 
    filter(!geo_area_name %in% no_country) |> 
    filter(goal %in% c(10,14,15))
```

```{r}
n_ctrs <- dat$geo_area_name |> unique() |> length() # number of countries

p <- dat |> 
    mutate(value = as.numeric(value), is_na = is.na(value)) |> 
    group_by(series_description, time_period, goal) |> 
    summarize(nas = sum(is_na) / n_ctrs) |> 
    mutate(series_description = str_trunc(series_description, width = 100, side = "right")) |> 
    ggplot(aes(time_period, series_description)) +
    geom_tile(aes(fill = nas)) +
    scale_fill_viridis_c() +
    facet_wrap(~goal, scales = "free", nrow = 3) +
    theme_light(base_size = 6)

ggsave(
    plot = p, filename = "UN_SDG_10-14-15.png", path = "figures/", device = "png",
    width = 7, height = 5, dpi = 400, bg = "white"
)

p + theme_light(base_size = 12)
```