### 2022-03-02 inequality-SDGs.Rmd

---
title: "SDGs"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

This notebooks explores synergies and trade-offs between inequality and other SDGs with the World Bank dataset on sustainable development goals at the national scale.

```{r}
dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
    janitor::clean_names()

dat
```
```{r}
key <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGSeries.csv"
) |> janitor::clean_names()

key |> pull(topic) |> unique()

countries <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()

countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)

```

```{r}
dat$indicator_name |> unique() |> length()
```
Extract environmental variables only

```{r}
env_topics <- key$topic |> str_subset("Environment")

env_vars <- key |> 
    filter(topic %in% env_topics) |> 
    pull(indicator_name) |> 
    unique()
```

```{r}
n_countries <- dat |> filter(country_code %in% countries_list) |> 
    pull(country_code) |> unique() |> length()

p <- dat |>
    filter(indicator_name %in% env_vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    group_by(indicator_name, year) |> 
    mutate(nas = is.na(value)) |> 
    summarise(missing = sum(nas) / n_countries) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6)

ggsave(
    plot = p, filename = "WB_environment_SDGs.png", path = "figures/", device = "png",
    width = 6, height = 4, dpi = 400, bg = "white"
)

p + theme_light(base_size = 8)
```
And the inequality variables

```{r}
ineq_topics <- c("Financial Sector: Access", )

key$topic |> unique()
```

## To-dos

Wait until Maike and Tong revise the selection of variables. Take the analysis from there. Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable. Once one take missing values into account, the number of variables and countries get reduced quickly.