### 2022-09-27 inequality-SDGs.Rmd

---
title: "SDGs"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

This notebooks explores synergies and trade-offs between inequality and other SDGs with the World Bank dataset on sustainable development goals at the national scale.

```{r}
dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
    janitor::clean_names()

dat
```
```{r}
# key <- read_csv(
#     file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGSeries.csv"
# ) |> janitor::clean_names()
key <- googlesheets4::read_sheet(
    "https://docs.google.com/spreadsheets/d/1T6rZ5T1qL4BPDL5oatMf3y8lLzhnIDvxIvkbJeBkdUc/edit?usp=sharing", sheet = 1
)

key <- key |> janitor::clean_names()
key |> pull(topic) |> unique()

countries <- read_csv(
    file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGCountry.csv") |> 
    janitor::clean_names()

countries_list <- countries |> filter(!is.na(currency_unit)) |> pull(country_code)

```

```{r}
dat$indicator_name |> unique() |> length()
```
Extract environmental variables only

```{r}
# env_topics <- key$topic |> str_subset("Environment")
# 
# env_vars <- key |> 
#     filter(topic %in% env_topics) |> 
#     pull(indicator_name) |> 
#     unique()
maike_vars <- key |> 
    filter(!is.na(maike_selection)) |> 
    pull(indicator_name) |> 
    unique()
```

```{r}
n_countries <- dat |> filter(country_code %in% countries_list) |> 
    pull(country_code) |> unique() |> length()

p <- dat |>
    filter(indicator_name %in% maike_vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    group_by(indicator_name, year) |> 
    mutate(nas = is.na(value)) |> 
    summarise(missing = sum(nas) / n_countries) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6)

# ggsave(
#     plot = p, filename = "WB_environment_SDGs.png", path = "figures/", device = "png",
#     width = 6, height = 4, dpi = 400, bg = "white"
# )

p + theme_light(base_size = 8)
```

```{r}
other_vars <- key |> 
    filter(!is.na(include)) |> 
    pull(indicator_name) |> 
    unique()

p <- dat |>
    filter(indicator_name %in% other_vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    group_by(indicator_name, year) |> 
    mutate(nas = is.na(value)) |> 
    summarise(missing = sum(nas) / n_countries) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    theme_light(base_size = 6)

p + theme_light(base_size = 8)
```

And the inequality variables

```{r}
# ineq_topics <- c("Financial Sector: Access", )

key$topic |> unique()
```

## To-dos

Wait until Maike and Tong revise the selection of variables. Take the analysis from there. Alternatively, select a couple of topics as example, just to explore how to deal with missing values and what is doable. Once one take missing values into account, the number of variables and countries get reduced quickly.

Combining both options:

```{r}
other_vars <- key |> 
    filter(!is.na(include) | !is.na(maike_selection)) |> 
    pull(indicator_name) |> 
    unique()

p <- dat |>
    filter(indicator_name %in% other_vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    group_by(indicator_name, year) |> 
    mutate(nas = is.na(value)) |> 
    summarise(missing = sum(nas) / n_countries) |> 
    ungroup() |> group_by(indicator_name) |> 
    summarize(mean_missing = mean(missing)) |> 
    filter(mean_missing < 0.3) |> 
    ggplot(aes(mean_missing, indicator_name)) +
    geom_point() +
    #scale_fill_viridis_c() +
    theme_light(base_size = 6)

p + theme_light(base_size = 8)

vars <- p$data$indicator_name

```
### Correlations

Probably need to log-transform cereal yield and forest area.

```{r}
dat |>
    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    select(-indicator_code) |> 
    pivot_wider(values_from = value, names_from = indicator_name) |> 
    GGally::ggpairs(
        columns = 4:12 ,
    lower = list(continuous = GGally::wrap("points", alpha = 0.5, size = 0.5)),
    diag = list(continuous = GGally::wrap("densityDiag", alpha = 0.2))
    # upper = list(continuous = GGally::wrap("cor", size = 2))
    ) + theme_light(base_size = 6)
```
Interpolate missing values

```{r}
df_dat <- dat |>
    filter(indicator_name %in% vars, country_code %in% countries_list) |> 
    select(-x35) |> 
    pivot_longer(cols = x1990:last_col(), values_to = "value", names_to = "year") |> 
    mutate(year = str_remove(year, "x"), year = as.numeric(year)) |> 
    select(-indicator_code) |> 
    pivot_wider(values_from = value, names_from = indicator_name) %>% 
    group_by(country_code) %>% 
    mutate(across(.cols = vars, .fns = na_interpolation, option = "spline"))

df_dat
```

## UN SDGs

```{r}
un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")
sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()

sdg <- sdg |> 
    left_join(
        select(un_countries, country = `Country or Area`, starts_with("ISO")),
        by = c("geo_area_name" = "country")) |> 
    filter(value != "-")|> 
    mutate(value = as.numeric(value))

sdg 
```

Why are there more than one obs per country per year per indicator???
```{r}
sdg |> 
    filter(`ISO-alpha3 Code` %in% countries_list) |> 
    group_by(geo_area_code, indicator, time_period) |> 
    summarize(n = n()) |> 
    filter(n>19) |>  # 19 is the max number of years
    arrange(desc(n))

sdg |> 
    filter(geo_area_code == 40, indicator == "5.4.1", time_period == 2009)

```
Answer: there are indicators that are calculated per age group, sex and rural / urban divide. They need to be processed to a single number per year. To discuss options:

- What dimension to use?: gender, age, rural / urban / all
- What summary statistic to use: mean, var, gini?

```{r}
n_countries2 <- sdg |> 
    filter(`ISO-alpha3 Code` %in% countries_list) |>
    pull(geo_area_name) |> unique() |> length() # number of countries, but here are also non-countries categories like continents, etc.

sdg |> 
    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
    filter(goal == 10) |> 
    group_by(indicator, indicator_name, year, iso3) |> 
    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
    pivot_wider(names_from = year, values_from = mean_val) |> 
    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
    mutate(nas = is.na(value)) |> 
    group_by(indicator_name, year) |> 
    summarise(missing = sum(nas) / n_countries2) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    labs(title = "SDG10: Inequality indicators") +
    theme_light(base_size = 8)

# sdg |> 
#     filter(series_description == "Proportion of voting rights of developing countries in international organizations, by organization (%)")

```

```{r}
sdg |> 
    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
    filter(goal == 14) |> 
    group_by(indicator, indicator_name, year, iso3) |> 
    summarize(mean_val = mean(value, na.rm = TRUE)) |> 
    pivot_wider(names_from = year, values_from = mean_val) |> 
    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
    mutate(nas = is.na(value)) |> 
    group_by(indicator_name, year) |> 
    summarise(missing = sum(nas) / n_countries2) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    labs(title = "SDG14: Life under water") +
    theme_light(base_size = 8)
```

```{r}
sdg |> 
    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
    filter(goal == 15) |> 
    group_by(indicator, indicator_name, year, iso3) |>
    summarize(mean_val = mean(value, na.rm = TRUE)) |>
    pivot_wider(names_from = year, values_from = mean_val) |> 
    pivot_longer(cols = starts_with("2"), names_to = "year", values_to = "value",
                 names_transform = list(year = as.numeric, value = as.numeric)) |> 
    mutate(nas = is.na(value)) |> 
    group_by(indicator_name, year) |> 
    summarise(missing = sum(nas) / n_countries2) |> 
     mutate(indicator_name = str_trunc(indicator_name, width = 30, side = "right")) |> 
    ggplot(aes(year, indicator_name)) +
    geom_tile(aes(fill = missing)) +
    scale_fill_viridis_c() +
    labs(title = "SDG15: Life on land") +
    theme_light(base_size = 8)

sdg|> 
    rename(indicator_name = series_description, year = time_period, iso3 =`ISO-alpha3 Code` ) |> 
    filter(iso3 %in% countries_list, !is.na(value)) |> # use only countries
    filter(goal == 15) |> 
    pull(indicator_name) |> unique()

```

## Notes:

Select the variables for which you have time series. Repeat the trajectories analysis. 
Do static analysis with all SDGs

```{r}
ctr <- sdg |> pull(geo_area_name) |> unique() |> length()
sdg <- sdg |> select(-c(age:type_of_speed))

sdg <- sdg |> 
    mutate(value = as.numeric(value)) |> 
    rename(year = time_period)
yrs <- sdg$year |> range() |> diff()
sdg <- sdg |> unique()

dat2 <- sdg |> 
    filter(goal == 10 | goal == 14 | goal == 15 ) |> 
    group_by(indicator, geo_area_code, year, geo_area_name, goal, target, series_code, series_description) |> 
    summarize(value = mean(value))

## If I add series description it duplicate entries, possibly due to misspellings
#test
dat2 |>
    unique() |>
    add_count() |>
    filter(n>1)

dat2 |> 
    add_count() |> 
    ungroup() |> 
    group_by(geo_area_name,  series_code) |> 
    summarize(yrs_cov = sum(n)) |> 
    filter(yrs_cov > 20)

sdg |> filter(series_code == "DC_ODA_BDVDL")
```