### 2023-06-14 wii-db.Rmd

---
title: "WIID database"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: inline
---

The World Income Inequality database has recently released new data for their
calculations of wealth and income inequality. Two resources are important:

- The [Code Dictionary](https://wid.world/codes-dictionary/): describes the 
rationale for new variables and how they are constructed. Currently there is over
1000 variables in the database
- The DINA guidelines is a more technical report (pdf) that explains the history
and legacy of the database, and describes the technical rationale for calculating
the different indicators. It is a good resource to understand limitations and 
caveats of each variable.

Both of these documents advise against using single number indicators (such as 
Gini) to describe inequality, and provide examples of when it fails. That is why
one needs the micro data, and why they make it available.

If we convert to PPP we need to use the last year of their convertion factors

> The WID contains two types of currency conversion factors (PPPs and market exchange rates) for three currencies (US dollars, euros and yuans). We provide long-term series for both of these indicators, but note that to convert between currencies in our benchmark approach, you only need to use the last value of each series: see Thomas Blanchet, Prices and currency conversions in WID.world, WID.world Technical Note Series n°2017/3 for details.

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(fs)
library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)
library(tictoc)
# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

Import data:

```{r}

fls <- dir_ls("~/Documents/Projects/DATA/WIID_World_Income_Inequality_DB/wid_all_data-20230323/")
# subseting countries only but there is subnational info for some nations (e.g. US)
fls <- fls |> str_subset(pattern = "WID_data_[:upper:]{2}\\.csv")

inq <- read_csv2(fls[44]) # 250 = US, the country with most data, 1093vas, 500k obs
inq                        # 44 Colombia, 215 Sweden, 211 South Africa

indicators <- map(fls, read_csv2) |> 
    map(function(x) x |> select(country,variable) |> unique()) |> 
    bind_rows()

indicators |> group_by(country) |> summarize(n=n()) |> arrange(desc(n)) # 277 countries
indicators |> group_by(variable) |> summarize(n = n() / 277) |> arrange(desc(n)) # 1377 variables

indicators |> 
    filter(str_detect(variable, "^g")) # g = gini, ptnic = pre-tax national income
```

Key variables:
- `ptninc` = pre-tax national income. It can be `g` for gini, `a` average, `b` inverted pareto-lorenz coefficient, `s` share, `r` ratio top 10 / bottom 50.
- `hweal` = net wealth, with same options as above.

```{r}
key_vars <- inq |> 
    pull(variable) |> 
    unique() |> 
    str_subset("(r|g|s)ptinc")

inq |> 
    filter(variable %in% key_vars) |># pull(percentile) |> unique()
    filter(percentile %in% c("p0p100", "p99p100")) |> 
    mutate(value = as.numeric(value)) |> 
    ggplot(aes(x = year, y = value)) +
    geom_line(aes(color = percentile), show.legend = FALSE) + 
    facet_wrap(~variable, scales = "free_y")
```

The ratio and gini are almost the same, at least for Colombia. Extract these variables for all files, all countries, all years:

```{r}
tic()
ineq_dat <- map(fls, read_csv2) |> 
    map(function(x) {
        x |> 
            filter(variable %in% key_vars) |># pull(percentile) |> unique()
            filter(percentile %in% c("p0p100", "p99p100")) |> 
            mutate(value = as.numeric(value), age = as.numeric(age))
        }) |> 
    bind_rows()
toc()

ineq_dat |> 
    filter(variable != "sptinc992j") |> 
    ggplot(aes(x = year, y = value)) +
    geom_line(aes(color = country), show.legend = FALSE) + 
    facet_wrap(~variable, scales = "free_y")
```

```{r}
ineq_dat |> 
    filter(variable != "sptinc992j") |> 
    pivot_wider(names_from = variable, values_from = value) |> 
    ggplot(aes(gptinc992j, rptinc992j)) +
    geom_point(aes(color = country), show.legend = FALSE, alpha = 0.5)

```

```{r}
library(naniar)
ineq_dat |> 
    filter(year >= 1980) |> 
    filter(variable != "sptinc992j") |> 
    ggplot(aes(year, value)) +
    geom_miss_point(aes(color = country), show.legend = FALSE, size = 0.1, alpha = 0.3) +
    geom_line(aes(color = country), show.legend = FALSE, linewidth = 0.1) +
    facet_wrap(~variable, scales = "free_y")
```

Ha! no missing values!!Add wealth variables and export for analysis.

```{r}
save(ineq_dat, file = "data/wii_small-data.RData")
```