### 2025-03-21 notes_retreat_2024.Rmd
subnational analysis draft

---
title: "Notes: retreat 2024"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Mechanisms & feedbacks session

1. Fairness and perceptions
2. Market concentration
3. Aspiration
    - conspicuous consumtpion
    - social norms
4. Cooperation
5. Environmental change
6. Environmental shocks

**Literature review workflow**

- map research: who, what, etc + key topics
- effects social inequality on environment

Problems:

- Levels, aggregation
- Measures
- direct vs indirect effects
- intertwinness
- positive feedbacks vs negative ones

Policy = P; Inequality = I; Enviornment = E

```{r}
library(DiagrammeR)
mermaid(
    "graph TB
    P --> I
    I --> E
    "
)

mermaid(
    "graph TB
    P <--> I
    I <--> E
    P <--> E
    "
)
```

Environmental justice: colonialims etc

Potential way of coding: