### 2025-03-21 subnational.Rmd

---
title: "Subnational inequality"
author: "Juan Rocha"
date: "Updated `r Sys.Date()`"
output:
  html_notebook:
    toc: yes
    toc_float: yes
    highlight: tango
    code_folding: hide
    df_print: paged
    theme: 
      bootswatch: cosmo
      code_font:
        google: Fira Code
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  cache = TRUE, warning = FALSE, message = FALSE, echo = TRUE,
  fig.width = 8, fig.height = 5
)

library(tidyverse)
library(naniar)
library(plotly)
library(patchwork)
library(fs)
library(tictoc)

# time series
library(tsibble)
library(fable)
library(imputeTS)
library(slider)

## spatial
library(terra)
library(tidyterra)
library(ggspatial)
library(stars)

# # clustering
# library(vegan)
# library(NbClust)
# library(clValid)
# library(mclust)

#theme_set(theme_light(base_size = 12))
```

## Inequality

Exploring subnational inequality data from Kummu et al. Load data:

```{r}

fls <- dir_ls("~/Documents/Projects/DATA/Kummu_inequality/transfer_294976_files_744c217c/")
dat <- rast(fls[2])
dat

dat <- vect(fls[5])
#terra::plot(dat$gini_disp_2021)

tmap::tm_polygons(dat)

ggplot(data = dat) +
    geom_spatvector(aes(fill = `2020`), color = "white", linewidth = 0.1) +
    scale_fill_viridis_c()
   # scale_fill_gradient2() 
   # 
dat |> as_tibble() |> 
    ggplot(aes(`1990`, `2021`)) +
    geom_point(aes(color = slope)) +
    geom_abline(slope = 1, intercept = 0) +
    scale_color_gradient2()
```

Time series

```{r}
dat |> 
    as_tibble() |> 
    pivot_longer(cols = `1990`:`2021`, names_to = "year", values_to = "Gini") |> 
    ggplot(aes(year, Gini)) +
    geom_line(aes(group = iso3), alpha = 0.5)
```

## Human footprint

```{r}
hfp <- dir_ls("/Users/juanrocha/Documents/Projects/DATA/HII_human_impact_index/first-gen/")

fp1 <- stars::read_stars(hfp[10])
fp1 <- rast(hfp[20])

terra::plot(fp1, type = "continuous", #col = map.pal("viridis", 100),
            range = c(-100,5000),
            col = c("#224f1a","#a3ff76","#feff6f","#a09568","#ffa802","#f7797c","#fb0102","#d87136","#a90086","#7a1ca5","#421137","#000000")
            )
#click(n=1)

# tic()
# m <- max(fp1, na.rm = TRUE)
# toc() # 496s

m

fp1[,10000,] |> range(na.rm= TRUE)
```

From Google Earth engine 
```{javascript eval=FALSE}
var hiiviz = {min: 5, max: 5000, palette: ["224f1a","a3ff76","feff6f","a09568","ffa802","f7797c","fb0102","d87136","a90086","7a1ca5","421137","000000"]
```

## Spatial joints

```{r}
#dat <- dat |> st_as_sf()
#plgns <- dat |> select(admID, iso3, geometry)

tic()
fp <- extract(x = fp1, y = dat, fun = mean, bind = TRUE, na.rm = TRUE)
toc() #
```