### 2025-05-06 paper/manuscript.Rmd
implementing comments paper, updating analysis SDG un

---
title: "Trade-offs between biosphere goals and inequality at the national scale"
output:
  pdf_document:
    dev: pdf
    keep_tex: yes
    toc: no
    latex_engine: xelatex
  rtf_document:
    toc: no
  word_document:
csl: /Users/juanrocha/Documents/styles/nature.csl
bibliography: refs.bib
citecolor: blue
documentclass: article
font: Arial
fontsize: 10pt
linkcolor: blue
urlcolor: blue
header-includes:
- \usepackage{dcolumn, rotating, longtable, lineno, float, array, tabularx, inputenc}
- \setmainfont{Arial}
- \renewcommand{\abstractname}{}
- \graphicspath{{figures/}}
- \usepackage[margin=2.5cm]{geometry}
- \linenumbers
- \usepackage{setspace}
---

```{r data, include=FALSE}
library(tidyverse)
library(here)

## word counts
## Abstract 150
## Intro 697
## Method 957
## Results 621
## Discussion 1434
## Conclusion 267
## Total main 3976 + 150 abstract
## 
## Targets: 
# 
# - NatComms: 5000w | 150w abstract | 4 figs or tables | Methods end 3000w | Refs
# 70max
# 
# - WorldDev: | 4500-10k words including all (refs, captions, tables, etc) | 300w
# abstract

# Notes:

# how to make it different from the trilemma?# 
# clearer story
# how much of the patterns hold constant if we remove indicators?
# sensitivity analysis for MFA / ordination and clustering.
```

# Outline

Target: EcolEcon?
/ People and Nature?
/ PNAS / ERL / NatSust?
[4000w]

Authors:

Introduction {800w}

-   Previous studies proposed the hypothesis of the trilemma

-   Why inequality and the biosphere / as opposed to other SDGs

-   Do the trade-offs hold at national scale with SDG data?

Methods {800w}:

-   SDG datasets at national scale collected by World Bank and United Nations
-   Selection of variables
-   Multiple Factor Analysis (PCA over time)
-   Clustering
-   Trajectories

Results {1000w}:

-   We confirm the results of Wu with three clusters of countries in two different SDGs datastets.
-   We confirm the existence of inequality traps or regime shifts in inequality.
-   We confirm the path dependency: the pathway from one regime to another is not random and offer insights on the type of policies to favour to achieve multiple goals. It also points to risks that need to be manage to avoid trade-offs.

Discussion {1000w}

-   Poor data quality overall.
    Biodiversity metrics either don't change or are focused on conservation.
    Thus it speaks of intentions not actual state variables.

-   Despite data challenges we confirm the existence of inequality regimes.

-   Some of the mechanisms by which inequality and the biosphere might be coupled occur at the scale of individuals or households.

-   Future research gaps: do the patterns hold at subnational scales?
    Do the patterns hold with biodiversity variables that actually reflects the state of biodiversity?
    Would the patterns be the same at local scales?
    Would they be different in human dominated systems such as cities?

Conclusion {400w}

SM

-   Figure with missing values from WB dataset: **improve resolution**

-   Figure with missing values from UN dataset: **improve resolution**

-   Correlograms

\pagebreak

\begin{abstract}

\end{abstract}

# Introduction

Countries around the world have committed to achieve 17 sustainable development goals as proposed by the United Nations.
The ambitious agenda is materialized in 169 targets and indicators, yet not all targets are monitored or properly measured, and not all countries report them, presenting significant data and knowledge gaps.
An open question in sustainability science is if these targets are achievable simultaneously?
Previous work on synergies and trade-offs between sustainable development goals (SDGs) suggest that both options are possible, yet this work is limited to expert elicitation and correlational studies.
It is unclear if there are system or structural constraints that limit a country's capacity to achieve particular SDGs, or if there are milestones that need to be unlocked before it can achieve the desired performance in a particular dimension of sustainability.
In other words, these correlations do not account for the path dependency and causal inference inherent on the process of development.

When the SDGs were adopted in 2015, work began to create a comprehensive database of SDG indicators [@schmidt-traub2017; @sachs2024], to be updated annually.
Based on these data, previous analyses found that goals associated with poverty alleviation, well-being, economic development, and innovation (e.g. SDGs 1, 3, 7, 8, 9) tended to be synergistic with other goals, while goals related to responsible consumption, climate action, natural resources, and cooperation showed the most trade-offs (e.g. SDGs 11, 12, 13, 14, 16, 17) [@pradhan2017; @s41599-019-0335-5].
Drawing upon relevant World Bank data, Lusseau and Mancini [@s41893-019-0231-4] find these patterns to be modulated by the overall income level of countries, with low-income countries showing synergies across all goals, while trade-offs start to appear in higher income categories.
More recently, [@s41467-023-44679-w] analysed transboundary SDG interactions and found that high income countries play a disproportionate role in influencing the achievement of SDGs in other countries.

Here we address the question of synergies and trade-offs by focusing on goals related to biosphere and inequality (SDGs 10, 14 and 15).
The motivation is threefold.
First, the Convention for Biological Diversity is currently negotiating and agreeing on the next set of goals and ambitions to bend the curve of biodiversity loss.
Understanding which countries are well positioned to achieve biodiversity goals or not will be necessary to design strategies that support countries with low capabilities, that address the dimensions of justice, and design targets with plausibility in mind.
Second, within-country inequality has been rising in the last decades even in high income countries, and as a consequence of the Covid pandemic, in between country inequality raised for first time in a generation.
These statistics set back progress on the inequality SDG by at least a decade, and undermines the mantra of leaving no one behind.
And third, recent conceptual and theoretical work has proposed mechanisms by which increase in inequality can impact the environment, and changes in the environment can feedback into further inequalities [@hamann2018; @dade2022; @leach2018].

Recent work also shows that a trilemma exists where countries struggle to achieve high prosperity, high environmental standards while reducing inequality [@Wu_2024].
Using data from environmental footprint, the gross national product, and the Gini coefficient time series, 53 countries were clustered and a typology of trajectories identified.
No country achieved the three goals simultaneously, and Latin American countries seem to have dynamics characterized by an inequality trap.
Countries trajectories on the trilemma space, however, suggest the hypothesis that social progress can be achieved without compromising the biosphere [@Wu_2024].

It remains an open question if these patterns hold across different datasets and scales.
If the hypothesis of the inequality trap is true, we should observe bimodal or multimodal distributions in inequality and environmental variables.
Each mode would correspond to an inequality regime, and the transitions probability of staying within one regime should be much higher than the probability of shifting regimes.
Here we address such questions by studying synergies and trade-offs between SDGs related to inequality and the biosphere.

-   Add a figure that describes the rationale of the methods and hypotheses

# Methods

```{r echo = FALSE, message=FALSE, warning=FALSE}
dat <- read_csv(file = "~/Documents/Projects/DATA/WorldBank/SDG_csv/SDGData.csv") |> 
    janitor::clean_names()

un_countries <- read_csv2(file = "~/Documents/Projects/DATA/SDGs_UNStats/UNSD — Methodology.csv")

# old UN dataset
# sdg <- read_csv(file = "~/Documents/Projects/DATA/SDGs_UNStats/20180817102044132_juan.rocha@su.se_data.csv") |> janitor::clean_names()

## new UN dataset:
fls <- dir_ls("data/sdgs/")

sdg <- list()
tic()
for (i in seq_along(fls)){
    load(fls[i])
    sdg[[i]] <- d
}
toc() # 21s

tic()
sdg <- sdg |> bind_rows() |> 
    unnest(c(goal, target, indicator))
toc() # 12.1s

# sdg <- sdg |> 
#     left_join(
#         select(un_countries, country = `Country or Area`, starts_with("ISO")),
#         by = c("geo_area_name" = "country")) |> 
#     filter(value != "-")|> 
#     mutate(value = as.numeric(value))

```

*Datasets:* To that end, we used the SDGs datasets made openly available by the World Bank (WB) and United Nations (UN).
The WB dataset offers `r dat$indicator_name |> unique() |> length()` indicators, with time series from 1990 to 2019 for `r dat$country_name |> unique() |> length()` countries or administrative areas (N = 2'013.791 observations).
The UN dataset offers time series from 1963 to 2025 for 17 SDGs, 168 targets, and 247 indicators, 687 time series, in 413 administrative areas (N = `r sdg |> nrow()` observations).
Despite their coverage, both datasets contain high proportion of missing values and the sample is unbalanced across countries and time, meaning some countries have better temporal coverage than others.
We focused subsequent analysis only on country level indicators for SDGs 5, 6, 10, 13, 14 and 15, related to inequality and the biosphere respectively.
We complemented the SDGs datasets with inequality data from the World Inequality database (WID), using their estimates of the ratio of pre-tax national income for working adults (population \> 20 years old) computed as the share of the top 10% over the share of the bottom 50% (`rptinc992j_p0p100`), the share of the 1% (`sptinc992j_p99p100`), and the Gini coefficient (`gptinc992j_p0p100`).
We also used their estimates for net wealth inequality computed as the share of the top 10% over the share of the bottom 50% (`rhweal992j_p0p100`), the share of the 1% (`shweal992j_p99p100`), and the Gini coefficient(`ghweal992j_p0p100`).

```{r echo = FALSE}
rm(dat, sdg, fls) # remove unnecessary objects
load(file = paste0(here(), "/data/cleaned_SDGs_2025.Rda")) 
# dat is the most recent UN dataset 2025, keys is the var names
load(file = paste0(here(), "/data/cleaned_sdg.RData"))
# `df_dat` is the data from World Bank, and `df_un_reduced` is the data from UN SDGs
df_dat <- df_dat |> ungroup()
df_un_reduced <- df_un_reduced |> ungroup()
```

*Variable selection:* We computed the proportion of missing values for all time series related to our initial selection of indicators (Figs S1, S2, S3).
We discarded indicators for which time series contained more than 30% of missing values, or less than 45 countries.
Missing values were then imputed using a cubic spline, leaving us with `r df_dat |> filter(!is.na(rptinc992j_p0p100)) |> pull(country_code) |> unique() |> length()` countries, `r df_dat |> pull(year) |> unique() |> length()` years of data across 9 indicators for the WB dataset; and `r keys |> filter(series %in% unique(dat$series)) |>  pull(series) |> unique() |> length()` series capturing `r keys |> filter(series %in% unique(dat$series)) |>  pull(indicator) |> unique() |> length()` indicators, `r dat |> pull(geoAreaName) |> unique() |> length()` countries over `r dat |> pull(timePeriodStart) |> unique() |> length()` years for the UN dataset.
The UN dataset was further reduced to 50 countries because the WID does not report inequality timeseries for Fiji.
Table X summarizes our selected variables, their units and available ranges.
A list of the countries analysed is presented in the supplementary information (SI).

```{r echo=FALSE}
# create table: source, series code, name, 
```

VOY AQUI

*Ordination:* We used multiple factor analysis (MFA) and principal component analysis (PCA) to reduce the dimensionality of the data and explore similarities and differences across countries.
MFA enables to specify the nested structure of our data and account for repeated observations of our variables over time.
We recover some of the qualitative results with PCA as robustness check, but these results are presented in the SI.
We also performed a clustering routine following the protocols by Charrad [] and Brooks [].
We tested over 10 clustering techniques and compared them across \>30 performance metrics to infer from the data what are the optimal numbers of clusters to fit and preferable algorithms.

*Analysis of trajectories:* With the results from the MFA we identified candidate variables where synergies or trade-offs are expected.
A trade-off in the reduced dimensional space occurs when improving on the direction of one indicator (e.g. reducing inequality) implies declining on the direction of another indicator.
Similarly, a synergy would be when increasing in one indicator seems to pull a country on the direction of improving in another indicator.
We extracted trajectories of the countries of analysis for some of these candidate variables.
If the hypothesis of inequality traps or inequality regimes is true, we would expect to find multimodal distributions in inequality indicators with respect to other variables; we expect the modes of these distributions to be correlated to the countries groups identified via clustering analysis; and we expect that most countries stay within their region of the parameter space and perhaps a few of them able to move between regimes.

```{r table1, include=TRUE, echo=FALSE, warning=FALSE, message=FALSE}

library(kableExtra)

un_vars <- df_un_reduced |> names() |> str_subset("^\\d{4}_") |> str_remove("\\d{4}_") |> unique()
un_vars <- un_vars[-c(7:9)]
wb_vars <- names(df_dat)[4:12]
iq_vars <-  names(df_dat)[c(14,16,17)]

tbl <- tibble(
    Source = c(rep("UN",length(un_vars)), rep("WB", length(wb_vars)), rep("WII",length(iq_vars))),
    Variable = c(un_vars, wb_vars, iq_vars)
    #Units = c(rep("Percentage", 4), "Thousands of hectares", "0-1")
)

tbl <- tbl |> mutate(
    Units = str_extract(Variable, "\\(.\\)$"),
    Units = str_remove_all(Units, "\\(|\\)")
) |> mutate(
        Variable = str_remove(Variable, " \\(%\\)"),
        Units = str_extract(Variable, "\\(.*\\)$"),
        Variable = str_remove(Variable, " \\(.*\\)$"),
        Units = str_remove_all(Units, "\\(|\\)"))

tbl$Units[1:4] <- "Percentage"
tbl$Units[6] <- "0 to 1"
tbl$Units[16:18] <- "0 to 1"
tbl$Variable[16:18] <- c(
    "Ratio of pre-tax national income for working adults (rptinc992j_p0p100)", "Gini coefficient (gptinc992j_p0p100)", "Share of the top 1% (sptinc992j_p99p100)"
)

knitr::kable(tbl, format = "latex", booktabs = TRUE, caption = "Summary of variables used") %>%
  kable_styling(latex_options = c("striped", "hold_position"), 
                full_width = TRUE, font_size = 8) %>%
  column_spec(1, width = "15mm") |> 
  column_spec(2, width = "10cm") 
```

# Results

Despite the differences in coverage with respect to countries, time, and indicators tracked, the ordination in both data sets results in three clusters of countries (Fig \ref{fig:wb}, Fig \ref{fig:un}).
For the WB data set, countries along the first principal component are differentiated by high levels of economic inequality, high energy intensity but low carbon emissions (negative values of Dim 1, green cluster), versus countries with relatively low gender inequality, high agricultural productivity, high carbon emissions, high urbanization and internet access (positive values in Dim 1, yellow and blue clusters).
Forest related variables have the lowest loading on the first two components and do not change over time, while access to internet or inclusion of women in leadership roles have the highest variability over time.

\begin{figure*}[ht]
\centering
\includegraphics[width = 7in, height = 4in]{fig_mfa_wb}
\caption{\textbf{Multiple factor analysis with World Bank data} The first 10 principal componets explain 94.8\% of the varation, the first two (A) explain 53\%. The 10 first components were used to cluster 151 countries resulting in three clusters (A). The correlation circle across explanatory variables is presented in (B) along their loadings on the first two components of the ordination.}
\label{fig:wb}
\end{figure*}

The UN dataset offers a similar ordination, where countries with low values along the first axes but higher values along the second axis have the highest inequalities (e.g Mexico, Fig \ref{fig:un}).
The inequality variables are highly correlated but also explain large amounts of the variance.
Contrary to the WB data, here forest related variables do show variability over time, but variables related to biodiversity loss (Red list index) do not change much over the time period of the data (2000-2015).
Interestingly, the ordination suggests that places with low inequality would have high Red list index.
Japan, Italy or Ireland are examples of countries on the region of the parameter space where inequality is relatively low but conservation metrics per country are relatively high.

\begin{figure*}[ht]
\centering
\includegraphics[width = 7in, height = 4in]{fig_UN_ordination}
\caption{\textbf{Multiple factor analysis with United Nations data}  }
\label{fig:un}
\end{figure*}

We find evidence of multi-modal distributions in inequality and environmental variables.
This finding supports the hypothesis of the existence of inequality regimes, particularly when inequality is measured as the share of the top 1% (Fig \ref{fig:bim}).
However, the lack of variability in environmental variables and smaller sample size in the UN data set prevent us to statistically derive modes in the distribution, except for the red list index (Fig \ref{fig:SM_UNbim}).

\begin{figure*}[ht]
\centering
\includegraphics[width = 7in, height = 4in]{bimodal}
\caption{\textbf{Inequality regimes} We find bimodal distributions for some dimensions of inequality against environmental factors in the WB dataset. Countries trajectories in A using the same cluster groups as in Fig \ref{fig:un}. Bimodal distributions are more common for the share of the top 1\% than other inequality variables (B). The y-axis has been rescaled to the range 0-1 to ease comparison. Supplementary figure \ref{fig:SM_UNbim} shows a similar plot for the UN data, however lack of variability in the environmental data prevents the identification of multiple modes in the distribution.}
\label{fig:bim}
\end{figure*}

# Discussion

-   Biodiversity variables are either limited to conservation, or resource production.
    Other biodiversity related values such as cultural values (intrinsic, recreational, spiritual) are not captured by SDG indicators.

-   Data gaps undermine the analysis of trajectories over time.
    For some of the variables where information exist, the statistics do not change (e.g. % forest area).
    Thus the indicators fail at capturing progress or lack of it towards the sustainable goals.

-   We find evidence for bistability in inequality, further supporting the hypothesis of inequality traps.
    However, data gaps and lack of variability in biodiversity indicators prevent us to further test features of regime shift dynamics such as hysteresis.
    Further work can look into this hypothesis by looking into other datasets or longer time series.

# Conclusion

Notes:

-   Paper figures:

    -   Plot with PCA selection (2-3 components?) and then change over time of the components. Do clustering on trayectories as well? -- and then add a map with the clusters for each dataset.

# References {.unnumbered}

::: {#refs}
:::

\pagebreak

# Supplementary Material {#sec:SM}

\renewcommand\thefigure{S\arabic{figure}}
\renewcommand\thetable{S\arabic{table}}
\setcounter{table}{0}
\setcounter{figure}{0}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 4in]{WB_environment_SDGs}
\caption{\textbf{World Bank environmental SDGs} }
\label{fig:SM_WB_env}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{WB_environment_SDGs_othervars}
\caption{\textbf{World Bank environmental SDGs additional variables} }
\label{fig:SM_WB_env2}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 6in]{WB_selected_vars}
\caption{\textbf{World Bank selected variables} }
\label{fig:SM_WB_select}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{WB_correlogram}
\caption{\textbf{Correlogram of World Bank selected variables} }
\label{fig:SM_WBcorr}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{UN_SDG_10-14-15}
\caption{\textbf{United Nations SDGs 10, 14 and 15 dataset} }
\label{fig:SM_UNdata}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{UN_selected_vars}
\caption{\textbf{UN selected variables} }
\label{fig:SM_UNselect}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{UN_reduced_vars}
\caption{\textbf{UN reduced variables} }
\label{fig:SM_UNred}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{UN_correlogram}
\caption{\textbf{Correlogram of United Nations selected variables} }
\label{fig:SM_UNcorr}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 6in, height = 5in]{sm_wb_ordinations}
\caption{\textbf{Variable importance in ordination on WB data} }
\label{fig:SM_WBord}
\end{figure*}
\begin{figure*}[ht]
\centering
\includegraphics[width = 4in, height = 3.5in]{bimodal_UN}
\caption{\textbf{Bimodality in UN data} The quality of the data does not allow for detection of bimodality in most UN variables. The lack of variation in biosphere related data does not allow the identification of areas in the parameter space where trajectories converge, except for the red list index, where we find multimodal distributions }
\label{fig:SM_UNbim}
\end{figure*}